Stack Safety as a Security Property

### 2020-07-09 main.tex
Pass on 7

This enforcement model is flexible enough to capture any kind of mechanism used
for dynamic enforcement of properties of interest, including the following
representative families of policies.
\apt{I don't really buy this, because we really depend on the ability to define
  variation and rollback. I think we should back off and say: this
  works well for micropolicies, and we conjecture that we can make it work
  for code-altering appraoches (pure software or capability).}

@@

\paragraph*{Micropolicies}
%
Micropolicies~\citep{pump_oakland2015} are a hardware-assisted tag-based
reference monitoring mechanism, which have been applied to stack safety by 
\citet{DBLP:conf/sp/RoesslerD18}.
Here the policy state consists of metadata tags (e.g., identifying stack frames)
attached to each value in memory or registers, and the policy step function
checks that each machine operation obeys a set of rules on tags
(e.g., that the current $\PCname$ tag matches the tag on the stack
location being accessed), and halts the machine if not.  Distinguishing
ordinary machine state and policy state is very natural in this setting.
%% Here the machine state is extended with tags paired with all components of the
%% system, and a monitor checks for each operation that the tags on the operands
%% obey a set of rules. One could treat tags directly as part of the machine
%% state, and build the rules into the step function. But local state
%% encapsulation properties as we frame them are not intuitive in their
%% interaction with tags or other additional enforcement state, so it is helpful
%% to separate tags for clarity. This also distinguishes a hardware fail-stop from
%% a software fail state.
%% \rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
%%   policy? }

\paragraph*{Capability machines}
%
%Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines like CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies, including 
the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}.  Capabilities tightly package pointers
(ordinary data) with base and bounds information (policy data),
and their use requires rewriting source code (e.g. to use capabilities in place
of ordinary pointers), so separating ordinary machine state from policy
state is delicate. \apt{But we conjecture it can be done...?}

%% Mapping the underlying protection
%% mechanisms to policy states and step function would be a delicate operation, as
%% those are tightly coupled to the hardware design. However, separating the
%% enforcement aspects of a concrete policy (such as StkTokens, where policy state
%% would encompass details like the abstract call stack, the capability sealing
%% scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
%% the machine would be relatively straightforward.  \bcp{Not sure that any of
%%   this is comprehensible... :-(}

\paragraph*{Software-only policies}
%
Some policies are implemented in software, by modifying code in the compiler to
(e.g.) perform bounds checking~\citep{NagarakatteZMZ09} or to insert stack
canaries~\citep{Cowan+98}. \apt{not sure what to say here}
%These approaches are represented as policies that
%accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
%being produced by the compiler.) The set of policy states is again the unit type
%and the step function constant. Property enforcement comes from the fact that
%only ill-formed states can induce a trace that violates the property.

%% \Paragraph*{Null policy}
%% %
%% The simplest possible policy is the neutral policy that carries no information
%% and does nothing. Its set of policy states is the unit type, and its policy step
%% function is the constant unit function, thereby allowing all calls to the machine
%% step function. Its initialization function is defined on all states. This
%% corresponds to the lifting of our machine model to the policy setting. Machines
%% that fulfill a property by construction will also fulfill that property with the
%% null policy.

@@

\paragraph*{Eager vs Lazy Policies}

The conservative policy we just described, while testable and enforceable, is
too slow to be of practical use. What we want, instead, is a somewhat more
permissive policy that moreover can be efficiently implemented. Most of the
performance overhead incurred by this kind of \emph{eager policy} steps from the
need to set stack activation tags as a frame is created and reset them as it is
destructed, as \citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
mitigate those costs, they propose optimizations like Lazy Tagging and Lazy
Clearing that relax and defer some of the checks of an eager policy, being more
permissive in their allowed operations and catching violations if and when their
effect becomes manifest.

These \emph{lazy policies} admit more efficient implementations, but they
deliberately allow violations of stack integrity. So it becomes important to
characterize what alternative property these policies ought to implement.

\sna{moved a lot of discussion to after property, along with Rob's
     comments on it}

@@

  A lazy policy permits some violations of integrity or confidentiality,
  justified because a truly harmful violation will be caught later.
  The natural question is, how do we characterize the protections provided
  by these policies once stack safety is broken? How does a harmful violation
  that must be caught later differ from a harmless one? The properties in this
  section use the observation model to draw this distinction.

@@

    \paragraph{Rolling Back to Idealized State}
      Observable integrity weakens eager integrity by allowing a returned
      state to differ from its call state at high integrity components,
      provided that the differences do not change the observable behavior
      of the rest of the program. It is not immediately clear what constitutes
      changing behavior; we construct an idealized return state that contains
      only the changes that the caller was allowed to make, and use the trace
      from that state as a baseline against which the real trace will be
      compared.

@@

      \[\begin{split}
        \propdef{ObsTraceInt} ~ \contour ~ \MPT \triangleq
        & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \\
        & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
          \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
      \end{split}\]

@@

      This requires restoring varied components to their original values
      in the trace from the variant state, since after return they will
      once again be legitimately accessible. To return to the example in
      \cref{fig:confex}, the varied final state \(\mach_r'\) still has
      -5 as a return address for {\tt g}, so the trace after that return will
      behave oddly in a way that doesn't reflect a confidentiality violation.
      So we cannot immediately compare traces from \(\mach_r\) and
      \(\mach_r'\). We must restore the varied values while preserving
      differences that have occurred during execution.

@@

      \[\begin{split}
        \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \triangleq
        & \head(\MPT) = \mpostate{\mach_e}{\pol_e}{\obs_e} \Rightarrow \\
        & \head(\machT) = (\mach_e',\obs_e') \Rightarrow \\
        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
        & \last(\machT) = (\mach_r',\obs_r') \Rightarrow \\
        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mach_r'' \hookrightarrow \machT' \Rightarrow \\
        & \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
        & \pi_o(\MPT') \lesssim \pi_o(\machT') \\
      \end{split}\]

@@

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. Let \(\callmap\) be a callmap; for each initial
      state and \(n\)-transition consisting of \(\mpstate{\mach_c}{\pol_c}\)
      and \(\mpstate{\mach_e}{\pol_e}\), take any variation \(\mach_e'
      \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
      and \(\mach_e' \hookrightarrow \machT | \ret{\mach_c}\). A system
      enjoys observable stack confidentiality if, for all such
      \(\MPT\) and \(\machT\), \(\propdef{ObsTraceConf} ~
      \mathit{Cof}(\mach_c,n) ~ \ret{\mach_c} ~ \MPT ~ \machT\).

@@

@@

    Observable properties allow us defer enforcement until a property
    violation would become visible, as in lazy policies. But by formalizing
    the connection to the eager policy, we can identify when lazy policies
    miss violations that can become visible. The basic lazy tag policy from
    \citet{DBLP:conf/sp/RoesslerD18} tags written memory with the depth in the
    stack at which it was written, and enforces that it must be read from the
    same depth. This allows violations of observable stack safety!
    Say we have functions {\tt foo}, {\tt bar} and {\tt baz}, where {\tt foo}
    calls {\tt bar}, and then after {\tt bar} returns {\tt foo} calls
    {\tt baz}. Since {\tt bar} and {\tt baz} are at the same stack depth, if
    {\tt bar} writes to {\tt foo}'s frame, {\tt baz} can then read the tainted
    write and use it to create an observable difference that violates both
    observable integrity and confidentiality without violating the policy.
    This can be fixed by generating distinct tags per-activation.

@@

  \paragraph{Relation to Stack Safety}

\rb{Implied by the property we started with?}

\rb{We can also test the lazy policy, even if less efficiently, and it should
  enforce an inductive property, which may be more amenable to proving --- this
  should be the connection to the following section.}

\rb{Explain and compare what each policy enforces.}

>>>>

\paragraph*{Null policy}
%
The simplest possible policy is the neutral policy that carries no information
and does nothing. Its set of policy states is the unit type, and its policy step
function is the constant unit function, thereby allowing all calls to the machine
step function. Its initialization function is defined on all states. This
corresponds to the lifting of our machine model to the policy setting. Machines
that fulfill a property by construction will also fulfill that property with the
null policy.

\paragraph*{Software-only policies}
%
Some policies are implemented in software, by modifying code in the compiler to
(e.g.) perform bounds checking \citep{NagarakatteZMZ09} or to insert stack
canaries \citep{Cowan+98}. These approaches are represented as policies that
accept\bcp{what does it mean for a policy to accept a state??} only initial machine states that are well-formed for the policy (due to
being produced by the compiler.) The set of policy states is again the unit type
and the step function constant. Property enforcement comes from the fact that
only ill-formed states can induce a trace that violates the property.

\paragraph*{Capability machines}
%
Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines like CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies involving, for
example, compartmentalization or the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}. Mapping the underlying protection
mechanisms to policy states and step function would be a delicate operation, as
those are tightly coupled to the hardware design. However, separating the
enforcement aspects of a concrete policy (such as StkTokens, where policy state
would encompass details like the abstract call stack, the capability sealing
scheme, etc. \rb{is this digging too deep? or on the contrary, say more?}) from
the machine would be relatively straightforward.  \bcp{Not sure that any of
  this is comprehensible... :-(}

@@

\paragraph*{Micropolicies}
%
Micropolicies are another major family of policies \citep{pump_oakland2015},
which are applied to stack safety in \citep{DBLP:conf/sp/RoesslerD18}.
Here the machine state is extended with tags paired with all components of the
system, and a monitor checks for each operation that the tags on the operands
obey a set of rules. One could treat tags directly as part of the machine
state, and build the rules into the step function. But local state
encapsulation properties as we frame them are not intuitive in their
interaction with tags or other additional enforcement state, so it is helpful
to separate tags for clarity. This also distinguishes a hardware fail-stop from
a software fail state.
\rb{TODO Similarly to above, should we consider tags as part of the machine or part of the
  policy? }

@@

@@

The conservative policy described in the previous section, while
testable and enforceable, is rather slow to be of practical use. What
we want, instead, is a somewhat more permissive policy that can also
be efficiently implemented. Most of the performance overhead incurred
stems from the need to set stack activation tags as a frame is created
and reset them as it is destructed, as
\citet{DBLP:conf/sp/RoesslerD18} confirm in their evaluation. To
mitigate those costs, they propose optimizations like Lazy Tagging and
Lazy Clearing that relax and defer some of the checks of an eager
policy, being more permissive in their allowed operations and catching
violations if and when their effect becomes manifest.
%
These \emph{lazy policies} admit more efficient implementations, but
they deliberately allow violations of stack integrity temporarily,
with the checks deferred until the point a violation truly becomes
harmful. The natural question then is, how do we characterize the
protections provided by these policies once stack safety is broken?
How does a harmful violation that must be caught later differ from a
harmless one? The properties in this section rely on our notion of
observations to draw this distinction.

@@

\paragraph{Rolling Back to Idealized State}

Observable integrity weakens eager integrity by allowing a returned
state to differ from its call state at high integrity components,
provided that the differences do not change the observable behavior of
the rest of the program. It is not immediately clear what constitutes
changing behavior: we construct an idealized return state that
contains only the changes that the caller was allowed to make, and use
the trace from that state as a baseline against which the real trace
will be compared.

@@

      \[\begin{split}
        \propdef{ObsTraceInt} ~ \contour ~ \MPT \triangleq
        & \mpostate{\mach_e}{\pol_e}{\obs_e} = \head(\MPT) \Rightarrow \\
        & \mpostate{\mach_r}{\pol_r}{\obs_r} = \last(\MPT) \Rightarrow \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT'
          \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
      \end{split}\]

@@

      This requires restoring varied components to their original
      values in the trace from the variant state, since after return
      they will once again be legitimately accessible. To return to
      the example in \cref{fig:confex}, the varied final state
      \(\mach_r'\) still has -5 as a return address for {\tt g}, so
      the trace after that return will behave oddly in a way that
      doesn't reflect a confidentiality violation.  That means we
      cannot immediately compare traces from \(\mach_r\) and
      \(\mach_r'\), but rather that we must restore the varied values
      while preserving differences that have occurred during
      execution.

@@

      \leo{This is starting to be too spread out as a definition. Pack a bit?}
      \[\begin{split}
        \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \triangleq
        & \head(\MPT) = \mpostate{\mach_e}{\pol_e}{\obs_e} \Rightarrow \\
        & \head(\machT) = (\mach_e',\obs_e') \Rightarrow \\
        & \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
        & \last(\machT) = (\mach_r',\obs_r') \Rightarrow \\
        & \mach_r'' = \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r') \Rightarrow\\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mach_r'' \hookrightarrow \machT' \Rightarrow \\
        & \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
        & \pi_o(\MPT') \lesssim \pi_o(\machT') \\
      \end{split}\]

@@

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. Let \(\callmap\) be a callmap; for each initial
      state and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
      to \(\mpstate{\mach_e}{\pol_e}\), take any variation \(\mach_e'
      \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
      and \(\mach_e' \hookrightarrow \machT | \ret{\mach_c}\). A system
      enjoys observable stack confidentiality if, for all such
      \(\MPT\) and \(\machT\), \(\propdef{ObsTraceConf} ~
      \mathit{Cof}(\mach_c,n) ~ \ret{\mach_c} ~ \MPT ~ \machT\).

@@

      \theorem Stack safety implies observable stack safety.

      \proof \leo{TODO: Whatever we say about the previous proof
        we have to say similar things here}

@@

    Observable properties allow us defer enforcement until a property
    violation would become visible, as in lazy policies. But by formalizing
    the connection to the eager policy, we can identify when lazy policies
    miss violations that can become visible. The basic lazy tag policy from
    \citet{DBLP:conf/sp/RoesslerD18} tags written memory with the depth in the
    stack at which it was written, and enforces that it must be read from the
    same depth. However, this allows violations of observable stack safety!
    Say we have functions {\tt foo}, {\tt bar} and {\tt baz}, where {\tt foo}
    calls {\tt bar}, and then after {\tt bar} returns {\tt foo} calls
    {\tt baz}. Since {\tt bar} and {\tt baz} are at the same stack depth, if
    {\tt bar} writes to {\tt foo}'s frame, {\tt baz} can then read the tainted
    write and use it to create an observable difference that violates both
    observable integrity and confidentiality without violating the policy.

@@
